Explicit B-spline regularization in diffeomorphic image registration

### 2013-12-09 paper.tex
BUG:  word change.
ENH:  Final edits, indeed.

Continuing from the work of \cite{christensen1996} and subsequent
exploration into the mathematical formalisms of diffeomorphisms
\citep[e.g.][]{dupuis1998}, the well-known Large Deformation
Diffeomorphic Metric Mapping (LDDMM) algorithm was proposed in
\cite{beg2005}.  In contrast to the mapping produced by
\cite{christensen1996}, LDDMM yields the geodesic solution in the
space of diffeomorphisms between two images. Since its introduction,
LDDMM has inspired much innovation in the image registration
literature.  Applying the log-Euclidean framework of
\cite{arsigny2006}, DARTEL (Diffeomorphic Anatomic Registration using
Exponential Lie algebra) uses a constant velocity field
parameterization to provide a fast, diffeomorphic alternative
\citep{ashburner2007}. Additionally, symmetrical considerations in the
velocity field parameterization are discussed in \cite{avants2008} in
the context of a cross correlation similarity metric.  By explicitly
symmetrizing the LDDMM formulation, this Symmetric Normalization (SyN)
minimizes the bias of the resulting transformation when selecting the
``fixed'' and ``moving'' images.  A greedy version of this algorithm
has proven successful in neuroimaging \citep{klein2009} and pulmonary
\citep{murphy2011} applications as well as in multi-atlas label fusion \citep{wang2012}.

@@

However, the two approaches produce characteristically
different solutions.  In addition to smoothing kernel differences,
Gaussian convolution tends to ``flatten'' the signal in contrast to an
approximation or fitting of the signal provided by the DMFFD approach.
Also, whereas Gaussian convolution operates entirely within
discretized space, the B-spline approximation routine constructs a
continuous object prior to any voxelwise reconstruction of the sampled
fields.  Interestingly, a similar comparison was made with respect to
Gaussian derivative estimation \citep{bourma2007}.  Although typically
estimated using truncated, discrete Gaussian convolution, an
alternative based on B-spline approximation demonstrated superior
performance with similar computational cost.
%Further elaboration on these differences
%are discussed in subsequent sections.

@@

\begin{figure}[htb]
  \centering
    \includegraphics[width=0.95\textwidth]{qualitative.jpg}
  \caption{Randomly selected axial slices showing qualitative differences between
  SyN and B-spline SyN.  Crosshairs marks regions of maximal Jacobian difference. 
  }
  \label{fig:qualitative}
\end{figure}

>>>>

Continuing from the work of \cite{christensen1996} and subsequent
exploration into the mathematical formalisms of diffeomorphisms
\citep[e.g.][]{dupuis1998}, the well-known Large Deformation
Diffeomorphic Metric Mapping (LDDMM) algorithm was proposed in
\cite{beg2005}.  In contrast to the mapping produced by
\cite{christensen1996}, LDDMM yields the geodesic solution in the
space of diffeomorphisms between two images. Since its introduction,
LDDMM has inspired much innovation in the image registration
literature.  Applying the log-Euclidean framework of
\cite{arsigny2006}, DARTEL (Diffeomorphic Anatomic Registration using
Exponential Lie algebra) uses a constant velocity field
parameterization to provide a fast, diffeomorphic alternative
\citep{ashburner2007}. Additionally, symmetrical considerations in the
velocity field parameterization are discussed in \cite{avants2008} in
the context of a cross correlation similarity metric.  By explicitly
symmetrizing the LDDMM formulation, this Symmetric Normalization (SyN)
approach minimizes the bias of the resulting transformation when selecting the
``fixed'' and ``moving'' images.  A greedy version of this algorithm
has proven successful in neuroimaging \citep{klein2009} and pulmonary
\citep{murphy2011} applications as well as in multi-atlas label fusion \citep{wang2012}.

@@

The two regularization approaches (i.e. Gaussian convolution vs. DMFFD), 
however,  produce characteristically
different solutions.  In addition to smoothing kernel differences,
Gaussian convolution tends to ``flatten'' the signal in contrast to an
approximation or fitting of the signal provided by the DMFFD approach.
Also, whereas Gaussian convolution operates entirely within
discretized space, the B-spline approximation routine constructs a
continuous object prior to any voxelwise reconstruction of the sampled
fields.  Interestingly, a similar comparison was made with respect to
Gaussian derivative estimation \citep{bourma2007}.  Although typically
estimated using truncated, discrete Gaussian convolution, an
alternative based on B-spline approximation demonstrated superior
performance with similar computational cost.
%Further elaboration on these differences
%are discussed in subsequent sections.

@@

\begin{figure}[htb]
  \centering
    \includegraphics[width=0.95\textwidth]{qualitative.jpg}
  \caption{Randomly selected axial slices showing qualitative differences between
  SyN and B-spline SyN.  Crosshairs indicate regions of maximal Jacobian difference. 
  }
  \label{fig:qualitative}
\end{figure}
